---
title: "extra_code"
output: html_document
date: "2024-01-22"
---


Extra GLMM models not being used 
```{r cars}

Deer models 
Using a negative binomial because of overdispersion in the Poisson model 
```{r cars}
#### Model 

# No deer in Seattle proper 

sewadeer <- all_dat %>% subset(city == "sewa") %>% 
  subset(Black_tailed_deer != 0)
# still lots in seattle data tho 


deermod <- glmmTMB(Black_tailed_deer ~ develop + EXPOSURE + EFFECT  + 
                  med_income +  
                  season + ( 1 | site), 
                data = all_dat, 
                offset = log(days.active),
                family = "nbinom2")
summary(deermod)
confint(deermod)

deermod2 <- glmmTMB(Black_tailed_deer ~ city + scale(develop) + scale(EXPOSURE) + scale(EFFECT)  + 
                     scale(med_income) + season + ( 1 | site), 
                data = all_dat, 
                offset = log(days.active),
                family = "nbinom2")

summary(deermod2)
confint(deermod2)
AICc(deermod2)
AICc(deermod)

#### Plot 
deernb <- sjPlot::plot_model(deermod2, title = "Black-tailed deer",
                             rm.terms = "season [Fall20,
                            Spring19, Spring20, Summer19, Summer20,
                            Winter19,Winter20]", 
              #               axis.title = "",
                   show.values = TRUE, show.p = TRUE) + 
  scale_color_sjplot("circus") + 
  scale_y_log10(limits = c(0.1, 100))

```


Squirrel models 
Using a negative binomial because of overdispersion in the Poisson model 
```{r cars}
#### Model 
 
squirmod <- glmmTMB(Eastern_gray_squirrel ~ develop + EXPOSURE + EFFECT  + 
                  med_income +  
                  season + ( 1 | site), 
                data = all_dat, 
                offset = log(days.active),
                family = "nbinom2")
summary(squirmod)
confint(squirmod)

squirmod2 <- glmmTMB(Eastern_gray_squirrel ~ city + scale(develop) + scale(EXPOSURE) + scale(EFFECT)  + 
                     scale(med_income) + season + ( 1 | site), 
                data = all_dat, 
                offset = log(days.active),
                family = "nbinom2")

summary(squirmod2)

#### Plot 
squirnb <- sjPlot::plot_model(squirmod2, title = "Eastern gray squirrel",
                             rm.terms = "season [Fall20,
                            Spring19, Spring20, Summer19, Summer20,
                            Winter19,Winter20]", 
         #                    axis.title = "",
                   show.values = TRUE, show.p = TRUE) + 
  scale_color_sjplot("circus") + 
  scale_y_log10(limits = c(0.1, 10))

```


Rabbit models 
Using a negative binomial because of overdispersion in the Poisson model 
```{r cars}
#### Model 
 
rabbmod <- glmmTMB(Eastern_cottontail_rabbit ~ develop + EXPOSURE + EFFECT  + 
                  med_income +  
                  season + ( 1 | site), 
                data = all_dat, 
                offset = log(days.active),
                family = "nbinom2")
summary(rabbmod)
confint(rabbmod)

rabbmod2 <- glmmTMB(Eastern_cottontail_rabbit ~ city + scale(develop) + scale(EXPOSURE) + scale(EFFECT)  + 
                     scale(med_income) + season + ( 1 | site), 
                data = all_dat, 
                offset = log(days.active),
                family = "nbinom2")

summary(rabbmod2)
confint(rabbmod2)
AICc(rabbmod)
AICc(rabbmod2)

#### Plot 
rabbnb <- sjPlot::plot_model(rabbmod2, title = "Eastern cottontail rabbit",
                             rm.terms = "season [Fall20,
                            Spring19, Spring20, Summer19, Summer20,
                            Winter19,Winter20]", 
            #                 axis.title = "",
                   show.values = TRUE, show.p = TRUE) + 
  scale_color_sjplot("circus") + 
  scale_y_log10(limits = c(0.1, 10))

```


Richness model 
Using a negative binomial because of overdispersion in the Poisson model 
```{r cars}
#### Model 
glimpse(all_dat_veg)
all_dat_veg$richness
richmod <- glmmTMB(richness ~ develop + EXPOSURE + EFFECT  + 
                  med_income +  ( 1 | site), 
                data = all_dat_veg, 
                family = "poisson")

richmod2 <- glmmTMB(richness ~ city + scale(develop) + scale(EXPOSURE) + 
                      scale(EFFECT)  + scale(med_income) + ( 1 | site), 
                data = all_dat_veg, 
                family = "poisson")
summary(richmod)
confint(richmod)

summary(richmod2)
AICc(richmod)
AICc(richmod2)

#### Plot 
richpois <- sjPlot::plot_model(richmod2, title = "Mammal community species richness",
                             rm.terms = "season [Fall20,
                            Spring19, Spring20, Summer19, Summer20,
                            Winter19,Winter20]", 
                 #           axis.title = "",
                   show.values = TRUE, show.p = TRUE) + 
  scale_color_sjplot("circus") + 
  scale_y_log10(limits = c(0.1, 10))

```


Pielou's evenness  
Using a gaussian 
```{r cars}
#### Model 

all_dat_veg$evenness
evenmod <- glmmTMB(evenness ~  develop + EXPOSURE + EFFECT  + 
                  med_income +  ( 1 | site),
                data = all_dat_veg, 
                family = "gaussian")
# does not converge

evenmod2 <- glmmTMB(evenness ~ city + scale(develop) + scale(EXPOSURE) + scale(EFFECT)  + scale(med_income) + ( 1 | site),
                data = all_dat_veg, 
                family = "gaussian")
summary(evenmod2)
confint(richmod)

AICc(richmod)
AICc(evenmod2)

#### Plot 
evengauss <- sjPlot::plot_model(evenmod2, title = "Mammal community species evenness",
                             rm.terms = "season [Fall20,
                            Spring19, Spring20, Summer19, Summer20,
                            Winter19,Winter20]", 
                            axis.title = "",
                   show.values = TRUE, show.p = TRUE) + 
  scale_color_sjplot("circus") + 
  scale_y_log10(limits = c(0.1, 10))

```
